\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-25 paper.qmd
fig 13 caption and layout

error_plot_pbmc <- ggplot(error_pbmc, 
                          aes(x = a1, 
                              y = log(MSE), 
                              colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  # geom_vline(xintercept = 15, linetype="solid", 
  #            color = "black", linewidth=0.8, alpha = 0.5) +
  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 3, 5, 7, 9, 11, 13, 15, 17, 19, 21, 24, 26)]) +
  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#ff7f00',
                              '#984ea3','#999999','#a65628','#f781bf')) +
  ylab("log(MSE)") +
  xlab(expression(a[1])) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))

@@

```{r}
#| fig-cap: "MSE from UMAP, tSNE, PHATE, PaCMAP, and TriMAP wwith diffferent (hyper-)parameter choices applied to PBMC3k dataset for different number of bin widths along the x-axis ($a_1$). Which is the most reasonable representation of the structures present in the PBMC3k data? The residual plot have a steep slope at the beginning, indicating that a smaller $a_1$ causes a smaller amount of error. Then, the slope gradually increases or level up, indicating that a higher number of $a_1$ generates a larger error. It was observed that for every $a_1$, the lowest error occurred with tSNE using the perplexity: $30$."
#| label: fig-pbmc-mse
#| fig-pos: H
#| out-height: 100%

free(error_plot_pbmc) / wrap_plots(nldr1c, nldr2c, nldr3c, nldr4c,
           nldr5c, nldr6c, nldr7c, nldr8c, ncol = 4)
```

>>>>

error_plot_pbmc <- ggplot(error_pbmc, 
                          aes(x = a1, 
                              y = log(MSE), 
                              colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  # geom_vline(xintercept = 15, linetype="solid", 
  #            color = "black", linewidth=0.8, alpha = 0.5) +
  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#ff7f00',
                              '#984ea3','#999999','#a65628','#f781bf')) +
  ylab("log(MSE)") +
  xlab(expression(paste("binwidth (", a[1], ")"))) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))

@@

```{r}
#| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in Figure 1) is the better representation using MSE for varying binwidth ($a_1$). Colour  used for the lines and points in the top plot and in the scatterplots represents NLDR layout (a-h). Layout f is universally poor. Layouts a, b, g, h that show large separations between clusters are universally suboptimal. Layout d with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases. The choice of best is between layouts c and e, that have small separations between oddly shaped clusters. Layout e is the best choice."
#| label: fig-pbmc-mse
#| fig-pos: H
#| out-height: 100%

free(error_plot_pbmc) + wrap_plots(nldr1c, nldr2c, nldr3c, nldr4c,
           nldr5c, nldr6c, nldr7c, nldr8c, ncol = 2)
```

### 2024-08-25 paper.tex
Updated Figure 14
fig 13 caption and layout

\caption{\label{fig-pbmc-mse}MSE from UMAP, tSNE, PHATE, PaCMAP, and
TriMAP wwith diffferent (hyper-)parameter choices applied to PBMC3k
dataset for different number of bin widths along the x-axis (\(a_1\)).
Which is the most reasonable representation of the structures present in
the PBMC3k data? The residual plot have a steep slope at the beginning,
indicating that a smaller \(a_1\) causes a smaller amount of error.
Then, the slope gradually increases or level up, indicating that a
higher number of \(a_1\) generates a larger error. It was observed that
for every \(a_1\), the lowest error occurred with tSNE using the
perplexity: \(30\).}

@@

\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc2-sc-1.pdf}

>>>>

\caption{\label{fig-pbmc-mse}Assessing which of the 8 NLDR layouts on
the PBMC3k data (shown in Figure 1) is the better representation using
MSE for varying binwidth (\(a_1\)). Colour used for the lines and points
in the top plot and in the scatterplots represents NLDR layout (a-h).
Layout f is universally poor. Layouts a, b, g, h that show large
separations between clusters are universally suboptimal. Layout d with
little separation performs well at tiny binwidth (where most points are
in their own bin) and poorly as binwidth increases. The choice of best
is between layouts c and e, that have small separations between oddly
shaped clusters. Layout e is the best choice.}

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc2-sc-1.pdf}
